Tropical Mathematics for the Lambda Calculus

### 2023-01-19 appVII.tex
commit

\begin{proposition}[universal property of the tensor product, cf.~\ref{Russo2007}]
The tensor product $M\otimes N$ is the codomain of the universal $\Lawv$-bimorphism $M\times N \to M\otimes N$.
\end{proposition}

@@

\begin{lemma}[cf.~\ref{Russo2007}]
\begin{itemize}
\item $M\otimes N\simeq N\otimes M$;

@@

The bialgebra modality $(!, \delta, \epsilon, \Delta, e, \nabla,u)$ is defined as follows:
\begin{itemize}
\item the comonad $(!,\delta,\epsilon)$ is given by:
\begin{align*}
\epsilon(f)& =  f(1) \\
%\OV{\mathrm{der}}& = \iota_{1}\\
\delta(f)(n) & = \bigvee\left\{ 
\iota_{n}([\iota_{i_{1}}(a_{1}), \dots , \iota_{i_{n}}(a_{n})]) \ \Big \vert \ 
a_{j}\in !_{i_{j}}X,  
a_{1}\cup \dots \cup a_{n}\leq f(i_{1}+\dots+ i_{n})
\right\}
\end{align*}

We have $\epsilon(\delta(\alpha))=\delta(\alpha)(1)=
\bigvee\{\iota_{n}(\alpha(n))\mid n\in \BB N\}=\alpha$, and 
\begin{align*}
!\epsilon(\delta(\alpha))(n)&=
\bigvee\{ [\epsilon(\alpha_{1}),\dots, \epsilon(\alpha_{n})]
\mid
[\alpha_{1},\dots, \alpha_{n}] \leq \delta(\alpha)(n)\}
 \\
 &=
 \bigvee\left \{ [\epsilon(\alpha_{1}),\dots, \epsilon(\alpha_{n})]
\ \Bigg \vert \ 
\alpha_{i}=\iota_{j_{i}}(a_{i}),\bigcup_{i}a_{i}\leq 
%
%[ \iota_{j_{1}^{i}}(a_{1}^{i}),\dots, \iota_{j_{s_{i}}^{i}}(a_{s_{i}}^{i})], 
%\left [\bigcup_{j}a_{j}^{1},\dots, \bigcup_{j}a_{j}^{n}\right] \leq 
\alpha(n)
\right \} \\
&=
 \bigvee\left \{ [\epsilon(\alpha_{1}),\dots, \epsilon(\alpha_{n})]
\ \Bigg \vert \ 
\alpha_{i}=\iota_{1}(x_{i}),[x_{1},\dots, x_{n}]\leq 
\alpha(n)
\right \} \\
&= \bigvee\left \{ [
x_{1},\dots, x_{n}] \ \Bigg \vert  \ 
[x_{1},\dots, x_{n}]\leq \alpha(n)
\right \} = \alpha(n)
 \end{align*}
%But since $\epsilon(\beta)=\beta(1)$, this means that we are reduced to consider multisets $\alpha_{j}^{i} $ of cardinality 1, i.e.~of the form $[a_{j}^{i}]$, so we have  
% $!\epsilon(\delta(\alpha))(n)=
%\bigvee\{ [a_{1}^{i},\dots, a_{n}^{i}]\mid \bigvee_{i}[a_{1}^{i},\dots, a_{n}^{i}]= \alpha(n)\}= \alpha(n)$.

@@

\item the cocommutative comonoid structure $(!X, \Delta, e)$ is  defined as follows:
\begin{align*}
\Delta(f)& = \bigvee
\left\{ \iota_{n}(a)\otimes \iota_{m}(b) \ \Big \vert \  
n,m\in \BB N, a\cup b\leq f(n+m)
\right\}\\
e(f)&= f(0)
\end{align*}

@@

\begin{align*}
(\Delta\otimes 1)\Big(\Delta(\alpha)\Big)&=
(\Delta\otimes 1)\left (\bigvee\left \{\iota_{n}(a)\otimes \iota_{m}(b)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}\right)\\
&= 
\bigvee\left \{\iota_{n_{1}}(a_{1})\otimes \iota_{n_{2}}(a_{2})\otimes \iota_{m}(b)\ \Big \vert \ a_{1}\cup a_{2}\cup b \leq \alpha(n_{1}+n_{2}+m)\right\}\\
&= 
\bigvee\left \{\iota_{n}(a)\otimes \iota_{m_{1}}(b_{1})\otimes \iota_{m_{2}}(b_{2})\ \Big \vert \ a\cup b_{1}\cup b_{2} \leq \alpha(n+m_{1}+m_{2})\right\}
\\
&= (1\otimes \Delta)\Big(\Delta(\alpha)\Big)
\end{align*}

\begin{align*}
(e\otimes 1)\Big(\Delta(\alpha)\Big) & = 
( e\otimes 1)\left (\bigvee\left \{\iota_{n}(a)\otimes \iota_{m}(b)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}\right)\\
&= 
\bigvee\left \{\iota_{m}(b)\ \Big \vert \ b \leq \alpha(m)\right\} =
\alpha
\end{align*}
and one can argue similarly for $(1\otimes e)(\Delta(\alpha))$.

\begin{align*}
\sigma(\Delta(\alpha)) & = 
\bigvee\left \{\iota_{m}(b)\otimes \iota_{n}(a)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}
\\
& = 
\bigvee\left \{\iota_{n}(a)\otimes \iota_{m}(b)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}
\\&= \Delta(\alpha)
\end{align*}

Finally, the commutation of $\Delta$ and $\delta$:
{\tiny
\begin{align*}
(\delta\otimes \delta)\Big(\Delta(\alpha)\Big) & =
(\delta\otimes\delta)\left(\bigvee
\left\{
\iota_{n}(a)\otimes \iota_{m}(b) \mid a\cup b \leq \alpha(n+m)
\right\}\right)\\
&=
\bigvee
\left \{\iota_{k_{1}}(
[\iota_{i_{1}}(a_{1}),\dots, \iota_{i_{k_{1}}}(a_{k_{1}})])\otimes
\iota_{k_{2}}(
[\iota_{j_{1}}(b_{1}),\dots, \iota_{j_{k_{2}}}(b_{k_{2}})]) \ \Bigg \vert \
\bigcup_{l}a_{l} \cup \bigcup_{l}b_{l} \leq \alpha\left (\sum_{l}i_{l}+\sum_{l}j_{l}\right)
\right\}
\\
&=
\Delta
\left(\bigvee\left \{
\iota_{n}(
[\iota_{i_{1}}(a_{1}),\dots, \iota_{i_{n}}(a_{n})]) \ \Bigg \vert \ 
\bigcup_{j}a_{j}\leq \alpha(\sum_{j}i_{j})
\right\}\right)
\\
& = \Delta(\delta(\alpha))
\end{align*}
}

@@

\begin{align*}
\nabla\left((\nabla\otimes 1)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\otimes \gamma_{k}\right)\right) & =
\nabla \left(
\bigvee_{k,n}
\iota_{n}\left( \bigvee_{u+v=n}\alpha_{k}(u)\cup \beta_{k}(v)
\right)
\otimes \gamma_{k}
\right)\\
&=
\bigvee_{k,n}\iota_{n}\left( \bigvee_{u+v+w=n}
\alpha_{k}(u)\cup \beta_{k}(v)\cup \gamma_{k}(w)\right) \\
& =
\nabla \left(
\bigvee_{n,k}\alpha_{k}\otimes 
\iota_{n}\left( \bigvee_{v+w=n}\beta_{k}(v)\cup \gamma_{k}(w)\right)
\right)\\
&= 
\nabla\left((1\otimes\nabla)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\otimes \gamma_{k}\right)\right) 
\end{align*}

\begin{align*}
\nabla\left(
(1\otimes u)\left(\bigvee_{k}\alpha_{k}\otimes 0\right)\right) & =
\nabla\left( \bigvee_{k}\alpha_{k}\otimes\iota_{0}(\epsilon)\right)
=
\bigvee_{n,k}\iota_{n}( \alpha_{k}(n) )
 =  \bigvee_{k}\alpha_{k} \otimes 0
\end{align*}
and one can argue similarly for $\nabla(u\otimes 1)(\alpha)=\alpha$.

\begin{align*}
\nabla\left (\sigma\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}
\right) \right) & = 
\nabla \left (
\bigvee_{k}\beta_{k}\otimes \alpha_{k}
\right)\\
&= 
\bigvee_{k,n, p+q=n} \iota_{n}(\beta_{k}(p)\cup \alpha_{k}(q))
\\
&= 
\bigvee_{k,n, p+q=n}\iota_{n}( \alpha_{k}(p)\cup \beta_{k}(q))
\\
&= \nabla\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
\end{align*}

@@

\begin{align*}
e\left (\nabla\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)\right)  & =
e\left(
\bigvee_{k,n, p+q=n}\iota_{n}(\alpha_{k}(p)\cup \beta_{k}(q))
\right)
\\
&= \bigvee_{k}\alpha_{k}(0)\cup\beta_{k}(0) \\
&= \bigvee_{k} 0\cup 0\\
&= h_{Q}\left(\bigvee_{k}\alpha_{k}(0)\otimes\beta_{k}(0))\right) = h\left( (e\otimes e)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)\right)
\end{align*}
where $h_{Q}:Q\otimes Q \to Q$ indicates the isomorphism $h(\epsilon\otimes \delta)=\epsilon+\delta$.

\begin{align*}
\Delta(u(\epsilon))) = \Delta(\iota_{0}(\epsilon)) &=
 \left(
\bigvee\left\{\iota_{m}(a)\otimes\iota_{n}(b) \ \Bigg \vert \ 
a\cup b \leq \iota_{0}(\epsilon)(n+m)\right\}\right)\\
&=\bigvee\left\{\iota_{0}(\epsilon_{1})\otimes\iota_{0}(\epsilon_{2}) \ \Bigg \vert \ 
\epsilon_{1}+\epsilon_{2} \leq \epsilon\right\}\\
&= (u\otimes u)(\epsilon)
\end{align*}
where we are identifying $\epsilon \in q$ with the equivalence class
$\epsilon \otimes 0 = \{ \langle \epsilon_{1},\epsilon_{2}\rangle\mid \epsilon_{1}+\epsilon_{2}=\epsilon\}\in 
Q\otimes Q$ via the isomorphism $h_{Q}$ described above.

@@

{\tiny
\begin{align*}
&(\nabla\otimes \nabla)  (1\otimes \sigma\otimes 1)(\Delta\otimes \Delta)\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right) \\
&=
(\nabla\otimes \nabla)(1\otimes \sigma\otimes 1)
\left(
\bigvee
\left\{
 \iota_{n}(a_{1})\otimes \iota_{m}(a_{2})\mid
a_{1}\cup a_{2}\leq \alpha_{k}(n+m)
\right\}\otimes
\left\{
 \iota_{n}(b_{1})\otimes \iota_{m}(b_{2})\mid
b_{1}\cup b_{2}\leq \beta_{k}(n+m)
\right\}
\right)\\
&=
(\nabla\otimes \nabla)(1\otimes \sigma\otimes 1)
\left(
\bigvee
\left\{
 \iota_{n}(a_{1})\otimes \iota_{m}(a_{2})
\otimes \iota_{n'}(b_{1})\otimes \iota_{m'}(b_{2})
\mid
a_{1}\cup a_{2}\leq \alpha_{k}(n+m),
b_{1}\cup b_{2}\leq \beta_{k}(n'+m')
\right\}
\right)\\
&=
(\nabla\otimes \nabla)\left(
\bigvee
\left\{
 \iota_{n}(a_{1})\otimes \iota_{n'}(b_{1})
\otimes \iota_{m}(a_{2})\otimes \iota_{m'}(b_{2})
\mid
a_{1}\cup a_{2}\leq \alpha_{k}(n+m),
b_{1}\cup b_{2}\leq \beta_{k}(n'+m')
\right\}
\right)\\
&=
\bigvee_{k,z,z'}\left\{
\left(
\iota_{z}\left( \bigvee_{p+q=z}
a_{1}\cup b_{1}\right)\right)\otimes
\left(\iota_{z'}\left( \bigvee_{p'+q'=z'}
a_{2}\cup b_{2}\right)\right)
\ \Bigg \vert \
a_{1}\cup a_{2} \leq \alpha_{k}(p+p'),
b_{1}\cup b_{2}\leq \beta_{k}(q+q')
\right\}
\\
&=\Delta\left(
\bigvee_{k,z}
\left\{
\iota_{z}\left( \bigvee_{p+q=z} \alpha_{k}(p)\cup \beta_{k}(q) \right)
\right\}
\right)
\\
&=
\Delta\left (\nabla\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right) \right)
\end{align*}
}

Finally, let us check the compatibility of $\epsilon$ and $\nabla$, which in $\Mod$ reads as
$ \epsilon\circ \nabla=(\epsilon \otimes e)\vee (e\otimes \epsilon)$:
\begin{align*}
\epsilon \left(\nabla\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)\right) & =
\epsilon\left (  \bigvee_{k,n,p+q=n}\iota_{n}(\alpha_{k}(p)\cup \beta_{k}(q))\right) \\
&=\left(\bigvee_{k}\alpha_{k}(1) \right) \vee \left(
\bigvee_{k}\beta_{k}(1)\right)
\\
&=
\left( (\epsilon\otimes e)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
\right)
\vee
\left( (e\otimes \epsilon)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
\right)
\\
&=
\Big((\epsilon\otimes e)\vee (e\otimes \epsilon)\Big)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
 \end{align*}

@@

For the last equation, we only check it on basic tensors:
\begin{align*}
& \nabla((\delta\otimes \eta)((1\otimes \nabla))(\Delta\otimes \eta)(\alpha\otimes x)) \\
&=\nabla((\delta\otimes \eta)((1\otimes \nabla))\left(
\bigvee
\{\iota_{n}(a)\otimes \iota_{m}(b)\otimes \iota_{1}(x)\mid a\cup b \leq \alpha(n+m)\}
\right)\\ 
&=\nabla\left ((\delta\otimes \eta)\left(
\bigvee
\{\iota_{n}(a)\otimes \iota_{m+1}(b\cup \{x\})\mid a\cup b \leq \alpha(n+m)\}\right)\right)
\\ 
&=\nabla\left (
\bigvee\left 
\{[\iota_{i_{1}}(c_{1}),\dots,\iota_{i_{r}}(c_{r})]
\otimes \iota_{1}( \iota_{m+1}(b\cup \{x\}))\ \Bigg \vert \ 
\bigcup_{j}c_{j}\cup b \leq \alpha\left (\sum_{j}i_{j}+m\right )\right\}\right)
\\ 
&=
\bigvee\left\{
[\iota_{i_{1}}(c_{1}),\dots,\iota_{i_{r}}(c_{r}), \iota_{m+1}(b\cup \{x\})] \ \Bigg \vert \
\bigcup_{j}c_{j}\cup b \leq \alpha\left (\sum_{j}i_{j}+m\right )\right\}
\\
&=
\delta\left(
\bigvee_{n}\iota_{n+1}(
\alpha(n)\cup \{x\})
\right)
\\
&=
\delta(\nabla(\alpha\otimes \iota_{1}(x))
\\
&=
\delta(\nabla((1\otimes \eta)(\alpha\otimes x))) 
\end{align*}

@@

{\tiny
\begin{align*}
\nabla((!f\otimes !g)(\Delta(\alpha))) & =
\nabla\left((!f\otimes !g)\left(
\bigvee\{\iota_{n}([x_{1},\dots, x_{n}])\otimes \iota_{m}([y_{1},\dots, y_{m}])\mid [x_{1},\dots, x_{n},y_{1},\dots y_{m}]\leq \alpha(n+m)\}
\right)\right)\\
&=
\nabla\left(
\bigvee\{\iota_{n}([f(x_{1}),\dots, f(x_{n})])\otimes \iota_{m}([g(y_{1}),\dots, g(y_{m})])\mid [x_{1},\dots, x_{n},y_{1},\dots y_{m}]\leq \alpha(n+m)\}
\right)\\
&=
\bigvee\{\iota_{n+m}([f(x_{1}),\dots, f(x_{n}),g(y_{1}),\dots, g(y_{m})])\mid [x_{1},\dots, x_{n},y_{1},\dots y_{m}]\leq \alpha(n+m)\}\\
&=
\bigvee_{n}\{ \iota_{n}([f(x_{1})\vee g(x_{1}),\dots, f(x_{n})\vee g(x_{1})]
\mid [x_{1},\dots, x_{n}]\leq \alpha(n)\}\\
&=
!(f\vee g)(\alpha)
\end{align*}
}
\begin{align*}
u(e(\alpha)) = u(\alpha(0))& =\iota_{0}(\alpha({0})) =
\iota_{0}(\alpha(0))\vee
\bigvee\{\iota_{n+1}([\underbrace{\bot,\dots, \bot}_{n+1}])\mid n\in \BB N \}=
(!\bot) (\alpha)
\end{align*}

>>>>

\begin{proposition}[universal property of the tensor product, cf.~\cite{Russo2007}]
The tensor product $M\otimes N$ is the codomain of the universal $\Lawv$-bimorphism $M\times N \to M\otimes N$.
\end{proposition}

@@

\begin{lemma}[cf.~\cite{Russo2007}]
\begin{itemize}
\item $M\otimes N\simeq N\otimes M$;

@@

The bialgebra modality $(!, \delta, \epsilon, \Delta, e, \nabla,u)$ is defined as follows:
\begin{itemize}
\item the comonad $(!,\delta,\epsilon)$ is given by:
{\small
\begin{align*}
\epsilon(f)& =  f(1) \\
%\OV{\mathrm{der}}& = \iota_{1}\\
\delta(f)(n) & = \bigvee\left\{ 
\iota_{n}([\iota_{i_{1}}(a_{1}), \dots , \iota_{i_{n}}(a_{n})]) \ \Big \vert \ 
a_{j}\in !_{i_{j}}X,  
a_{1}\cup \dots \cup a_{n}\leq f(i_{1}+\dots+ i_{n})
\right\}
\end{align*}
}

We have $\epsilon(\delta(\alpha))=\delta(\alpha)(1)=
\bigvee\{\iota_{n}(\alpha(n))\mid n\in \BB N\}=\alpha$, and 
{\small
\begin{align*}
!\epsilon(\delta(\alpha))(n)&=
\bigvee\{ [\epsilon(\alpha_{1}),\dots, \epsilon(\alpha_{n})]
\mid
[\alpha_{1},\dots, \alpha_{n}] \leq \delta(\alpha)(n)\}
 \\
 &=
 \bigvee\left \{ [\epsilon(\alpha_{1}),\dots, \epsilon(\alpha_{n})]
\ \Bigg \vert \ 
\alpha_{i}=\iota_{j_{i}}(a_{i}),\bigcup_{i}a_{i}\leq 
%
%[ \iota_{j_{1}^{i}}(a_{1}^{i}),\dots, \iota_{j_{s_{i}}^{i}}(a_{s_{i}}^{i})], 
%\left [\bigcup_{j}a_{j}^{1},\dots, \bigcup_{j}a_{j}^{n}\right] \leq 
\alpha(n)
\right \} \\
&=
 \bigvee\left \{ [\epsilon(\alpha_{1}),\dots, \epsilon(\alpha_{n})]
\ \Bigg \vert \ 
\alpha_{i}=\iota_{1}(x_{i}),[x_{1},\dots, x_{n}]\leq 
\alpha(n)
\right \} \\
&= \bigvee\left \{ [
x_{1},\dots, x_{n}] \ \Bigg \vert  \ 
[x_{1},\dots, x_{n}]\leq \alpha(n)
\right \} = \alpha(n)
 \end{align*}
 }
%But since $\epsilon(\beta)=\beta(1)$, this means that we are reduced to consider multisets $\alpha_{j}^{i} $ of cardinality 1, i.e.~of the form $[a_{j}^{i}]$, so we have  
% $!\epsilon(\delta(\alpha))(n)=
%\bigvee\{ [a_{1}^{i},\dots, a_{n}^{i}]\mid \bigvee_{i}[a_{1}^{i},\dots, a_{n}^{i}]= \alpha(n)\}= \alpha(n)$.

@@

\item the cocommutative comonoid structure $(!X, \Delta, e)$ is  defined as follows:
{\small
\begin{align*}
\Delta(f)& = \bigvee
\left\{ \iota_{n}(a)\otimes \iota_{m}(b) \ \Big \vert \  
n,m\in \BB N, a\cup b\leq f(n+m)
\right\}\\
e(f)&= f(0)
\end{align*}
}

@@

{\small
\begin{align*}
(\Delta\otimes 1)\Big(\Delta(\alpha)\Big)&=
(\Delta\otimes 1)\left (\bigvee\left \{\iota_{n}(a)\otimes \iota_{m}(b)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}\right)\\
&= 
\bigvee\left \{\iota_{n_{1}}(a_{1})\otimes \iota_{n_{2}}(a_{2})\otimes \iota_{m}(b)\ \Big \vert \ a_{1}\cup a_{2}\cup b \leq \alpha(n_{1}+n_{2}+m)\right\}\\
&= 
\bigvee\left \{\iota_{n}(a)\otimes \iota_{m_{1}}(b_{1})\otimes \iota_{m_{2}}(b_{2})\ \Big \vert \ a\cup b_{1}\cup b_{2} \leq \alpha(n+m_{1}+m_{2})\right\}
\\
&= (1\otimes \Delta)\Big(\Delta(\alpha)\Big)
\end{align*}
}

{\small
\begin{align*}
(e\otimes 1)\Big(\Delta(\alpha)\Big) & = 
( e\otimes 1)\left (\bigvee\left \{\iota_{n}(a)\otimes \iota_{m}(b)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}\right)\\
&= 
\bigvee\left \{\iota_{m}(b)\ \Big \vert \ b \leq \alpha(m)\right\} =
\alpha
\end{align*}
}
and one can argue similarly for $(1\otimes e)(\Delta(\alpha))$.

{\small
\begin{align*}
\sigma(\Delta(\alpha)) & = 
\bigvee\left \{\iota_{m}(b)\otimes \iota_{n}(a)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}
\\
& = 
\bigvee\left \{\iota_{n}(a)\otimes \iota_{m}(b)\ \Big \vert \ a\cup b \leq \alpha(n+m)\right\}
\\&= \Delta(\alpha)
\end{align*}
}

Finally, the commutation of $\Delta$ and $\delta$:
{\small
\begin{align*}
(\delta\otimes \delta)\Big(\Delta(\alpha)\Big) & =
(\delta\otimes\delta)\left(\bigvee
\left\{
\iota_{n}(a)\otimes \iota_{m}(b) \mid a\cup b \leq \alpha(n+m)
\right\}\right)\\
&=
\bigvee
\left \{\iota_{k_{1}}(
[\iota_{i_{1}}(a_{1}),\dots, \iota_{i_{k_{1}}}(a_{k_{1}})])\otimes
\iota_{k_{2}}(
[\iota_{j_{1}}(b_{1}),\dots, \iota_{j_{k_{2}}}(b_{k_{2}})]) \ \Bigg \vert \
\bigcup_{l}a_{l} \cup \bigcup_{l}b_{l} \leq \alpha\left (\sum_{l}i_{l}+\sum_{l}j_{l}\right)
\right\}
\\
&=
\Delta
\left(\bigvee\left \{
\iota_{n}(
[\iota_{i_{1}}(a_{1}),\dots, \iota_{i_{n}}(a_{n})]) \ \Bigg \vert \ 
\bigcup_{j}a_{j}\leq \alpha(\sum_{j}i_{j})
\right\}\right)
\\
& = \Delta(\delta(\alpha))
\end{align*}
}

@@

{\small
\begin{align*}
\nabla\left((\nabla\otimes 1)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\otimes \gamma_{k}\right)\right) & =
\nabla \left(
\bigvee_{k,n}
\iota_{n}\left( \bigvee_{u+v=n}\alpha_{k}(u)\cup \beta_{k}(v)
\right)
\otimes \gamma_{k}
\right)\\
&=
\bigvee_{k,n}\iota_{n}\left( \bigvee_{u+v+w=n}
\alpha_{k}(u)\cup \beta_{k}(v)\cup \gamma_{k}(w)\right) \\
& =
\nabla \left(
\bigvee_{n,k}\alpha_{k}\otimes 
\iota_{n}\left( \bigvee_{v+w=n}\beta_{k}(v)\cup \gamma_{k}(w)\right)
\right)\\
&= 
\nabla\left((1\otimes\nabla)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\otimes \gamma_{k}\right)\right) 
\end{align*}
}

{\small
\begin{align*}
\nabla\left(
(1\otimes u)\left(\bigvee_{k}\alpha_{k}\otimes 0\right)\right) & =
\nabla\left( \bigvee_{k}\alpha_{k}\otimes\iota_{0}(\epsilon)\right)
=
\bigvee_{n,k}\iota_{n}( \alpha_{k}(n) )
 =  \bigvee_{k}\alpha_{k} \otimes 0
\end{align*}
}
and one can argue similarly for $\nabla(u\otimes 1)(\alpha)=\alpha$.

{\small
\begin{align*}
\nabla\left (\sigma\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}
\right) \right) & = 
\nabla \left (
\bigvee_{k}\beta_{k}\otimes \alpha_{k}
\right)\\
&= 
\bigvee_{k,n, p+q=n} \iota_{n}(\beta_{k}(p)\cup \alpha_{k}(q))
\\
&= 
\bigvee_{k,n, p+q=n}\iota_{n}( \alpha_{k}(p)\cup \beta_{k}(q))
\\
&= \nabla\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
\end{align*}
}

@@

{\small
\begin{align*}
e\left (\nabla\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)\right)  & =
e\left(
\bigvee_{k,n, p+q=n}\iota_{n}(\alpha_{k}(p)\cup \beta_{k}(q))
\right)
\\
&= \bigvee_{k}\alpha_{k}(0)\cup\beta_{k}(0) \\
&= \bigvee_{k} 0\cup 0\\
&= h\left(\bigvee_{k}\alpha_{k}(0)\otimes\beta_{k}(0))\right) = h\left( (e\otimes e)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)\right)
\end{align*}
}
where $h:\Lawv\otimes\Lawv\to \Lawv$ indicates the isomorphism $h(\epsilon\otimes \delta)=\epsilon+\delta$.

{\small
\begin{align*}
\Delta(u(\epsilon))) = \Delta(\iota_{0}(\epsilon)) &=
 \left(
\bigvee\left\{\iota_{m}(a)\otimes\iota_{n}(b) \ \Bigg \vert \ 
a\cup b \leq \iota_{0}(\epsilon)(n+m)\right\}\right)\\
&=\bigvee\left\{\iota_{0}(\epsilon_{1})\otimes\iota_{0}(\epsilon_{2}) \ \Bigg \vert \ 
\epsilon_{1}+\epsilon_{2} \leq \epsilon\right\}\\
&= (u\otimes u)(\epsilon)
\end{align*}
}
where we are identifying $\epsilon \in q$ with the equivalence class
$\epsilon \otimes 0 = \{ \langle \epsilon_{1},\epsilon_{2}\rangle\mid \epsilon_{1}+\epsilon_{2}=\epsilon\}\in 
\Lawv\otimes\Lawv$ via the isomorphism $h$ described above.

@@

{\small
\begin{align*}
&(\nabla\otimes \nabla)  (1\otimes \sigma\otimes 1)(\Delta\otimes \Delta)\left (\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right) \\
&=
(\nabla\otimes \nabla)(1\otimes \sigma\otimes 1)
\left(
\bigvee
\left\{
 \iota_{n}(a_{1})\otimes \iota_{m}(a_{2})\mid
a_{1}\cup a_{2}\leq \alpha_{k}(n+m)
\right\}\otimes
\left\{
 \iota_{n}(b_{1})\otimes \iota_{m}(b_{2})\mid
b_{1}\cup b_{2}\leq \beta_{k}(n+m)
\right\}
\right)\\
&=
(\nabla\otimes \nabla)(1\otimes \sigma\otimes 1)
\left(
\bigvee
\left\{
 \iota_{n}(a_{1})\otimes \iota_{m}(a_{2})
\otimes \iota_{n'}(b_{1})\otimes \iota_{m'}(b_{2})
\mid
a_{1}\cup a_{2}\leq \alpha_{k}(n+m),
b_{1}\cup b_{2}\leq \beta_{k}(n'+m')
\right\}
\right)\\
&=
(\nabla\otimes \nabla)\left(
\bigvee
\left\{
 \iota_{n}(a_{1})\otimes \iota_{n'}(b_{1})
\otimes \iota_{m}(a_{2})\otimes \iota_{m'}(b_{2})
\mid
a_{1}\cup a_{2}\leq \alpha_{k}(n+m),
b_{1}\cup b_{2}\leq \beta_{k}(n'+m')
\right\}
\right)\\
&=
\bigvee_{k,z,z'}\left\{
\left(
\iota_{z}\left( \bigvee_{p+q=z}
a_{1}\cup b_{1}\right)\right)\otimes
\left(\iota_{z'}\left( \bigvee_{p'+q'=z'}
a_{2}\cup b_{2}\right)\right)
\ \Bigg \vert \
a_{1}\cup a_{2} \leq \alpha_{k}(p+p'),
b_{1}\cup b_{2}\leq \beta_{k}(q+q')
\right\}
\\
&=\Delta\left(
\bigvee_{k,z}
\left\{
\iota_{z}\left( \bigvee_{p+q=z} \alpha_{k}(p)\cup \beta_{k}(q) \right)
\right\}
\right)
\\
&=
\Delta\left (\nabla\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right) \right)
\end{align*}
}

Finally, let us check the compatibility of $\epsilon$ and $\nabla$, which in $\Mod$ reads as
$ \epsilon\circ \nabla=(\epsilon \otimes e)\vee (e\otimes \epsilon)$:
{\small\begin{align*}
\epsilon \left(\nabla\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)\right) & =
\epsilon\left (  \bigvee_{k,n,p+q=n}\iota_{n}(\alpha_{k}(p)\cup \beta_{k}(q))\right) \\
&=\left(\bigvee_{k}\alpha_{k}(1) \right) \vee \left(
\bigvee_{k}\beta_{k}(1)\right)
\\
&=
\left( (\epsilon\otimes e)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
\right)
\vee
\left( (e\otimes \epsilon)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
\right)
\\
&=
\Big((\epsilon\otimes e)\vee (e\otimes \epsilon)\Big)\left(\bigvee_{k}\alpha_{k}\otimes \beta_{k}\right)
 \end{align*}
}

@@

For the last equation, we only check it on basic tensors:
{\small
\begin{align*}
& \nabla((\delta\otimes \eta)((1\otimes \nabla))(\Delta\otimes \eta)(\alpha\otimes x)) \\
&=\nabla((\delta\otimes \eta)((1\otimes \nabla))\left(
\bigvee
\{\iota_{n}(a)\otimes \iota_{m}(b)\otimes \iota_{1}(x)\mid a\cup b \leq \alpha(n+m)\}
\right)\\ 
&=\nabla\left ((\delta\otimes \eta)\left(
\bigvee
\{\iota_{n}(a)\otimes \iota_{m+1}(b\cup \{x\})\mid a\cup b \leq \alpha(n+m)\}\right)\right)
\\ 
&=\nabla\left (
\bigvee\left 
\{[\iota_{i_{1}}(c_{1}),\dots,\iota_{i_{r}}(c_{r})]
\otimes \iota_{1}( \iota_{m+1}(b\cup \{x\}))\ \Bigg \vert \ 
\bigcup_{j}c_{j}\cup b \leq \alpha\left (\sum_{j}i_{j}+m\right )\right\}\right)
\\ 
&=
\bigvee\left\{
[\iota_{i_{1}}(c_{1}),\dots,\iota_{i_{r}}(c_{r}), \iota_{m+1}(b\cup \{x\})] \ \Bigg \vert \
\bigcup_{j}c_{j}\cup b \leq \alpha\left (\sum_{j}i_{j}+m\right )\right\}
\\
&=
\delta\left(
\bigvee_{n}\iota_{n+1}(
\alpha(n)\cup \{x\})
\right)
\\
&=
\delta(\nabla(\alpha\otimes \iota_{1}(x))
\\
&=
\delta(\nabla((1\otimes \eta)(\alpha\otimes x))) 
\end{align*}
}

@@

{\small
\begin{align*}
\nabla((!f\otimes !g)(\Delta(\alpha))) & =
\nabla\left((!f\otimes !g)\left(
\bigvee\{\iota_{n}([x_{1},\dots, x_{n}])\otimes \iota_{m}([y_{1},\dots, y_{m}])\mid [x_{1},\dots, x_{n},y_{1},\dots y_{m}]\leq \alpha(n+m)\}
\right)\right)\\
&=
\nabla\left(
\bigvee\{\iota_{n}([f(x_{1}),\dots, f(x_{n})])\otimes \iota_{m}([g(y_{1}),\dots, g(y_{m})])\mid [x_{1},\dots, x_{n},y_{1},\dots y_{m}]\leq \alpha(n+m)\}
\right)\\
&=
\bigvee\{\iota_{n+m}([f(x_{1}),\dots, f(x_{n}),g(y_{1}),\dots, g(y_{m})])\mid [x_{1},\dots, x_{n},y_{1},\dots y_{m}]\leq \alpha(n+m)\}\\
&=
\bigvee_{n}\{ \iota_{n}([f(x_{1})\vee g(x_{1}),\dots, f(x_{n})\vee g(x_{1})]
\mid [x_{1},\dots, x_{n}]\leq \alpha(n)\}\\
&=
!(f\vee g)(\alpha)
\end{align*}
}
\begin{align*}
u(e(\alpha)) = u(\alpha(0))& =\iota_{0}(\alpha({0})) =
\iota_{0}(\alpha(0))\vee
\bigvee\{\iota_{n+1}([\underbrace{\bot,\dots, \bot}_{n+1}])\mid n\in \BB N \}=
(!\bot) (\alpha)
\end{align*}
